---
title: "Stats 20: Homework 2"
author: "Brandon Liu -- UID: 306-375-425"
date: "`2025-01-28`"
output:
  pdf_document:
    toc: true
    toc_depth: '4'
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
    code_folding: show
---

\newpage
## Question 1

```{r Question 1 Background}
mixed1 <- c(TRUE, FALSE, FALSE, TRUE, 4, 0, 3)
mixed1
mixed2 <- c(TRUE, FALSE, FALSE, TRUE, 4, 0, "3")
mixed2
mixed3 <- c(c(TRUE, FALSE, FALSE, TRUE, 4, 0), "3")
mixed3
```

### (a)
mixed 2 outputted all inputs in the form of character vectors since the concatenation function was concatenating logical vectors, numerical vectors, and character vectors, and character vectors supersede all other modes.

mixed3 outputted True/False logical functions in their numerical vector equivalents of 1s and 0s since the first concatenation function was concatenating the logical vectors and numerical vectors, and numerical vectors supersede logical vectors. With the second concatenation function, 3 was inputted as a character vector and was outputted as such.

### (b)
<!-- ENTER YOUR ANSWER BELOW, IT CAN BE JUST THE COMMAND. NO NEED TO STORE AN OBJECT -->
```{r Question 1b}
c(c("TRUE", "FALSE"), c(FALSE, TRUE, 4, 0, 3))
```

\newpage
## Question 2

### (a)
```{r Question 2a}
as.numeric(mixed2)
```

```{r Question 2ai}
as.numeric(mixed3)
```
as.numeric(mixed2) converts all character vector outputs to numeric vectors. NAs (not availables) are returned because the character vectors of "TRUE" and "FALSE" cannot be converted to a numeric vector. If they were in their regular logical vector form of TRUE and FALSE, then they would convert to 1s and 0s.

as.numeric(mixed3) converts the TRUE and FALSE inputs to numeric vectors due to the initial concatenation function. When running the as.numeric() function, the TRUEs and FALSEs outputted as 1s and 0s since they were already set to numeric vector form.

### (b)
```{r Question 2b}
as.logical(mixed2)
as.logical(mixed3)
```
as.logical(mixed2) converts all character vectors to logical vectors. Since  "TRUE" and "FALSE" correspond to TRUE and FALSE logical vectors, these inputs printed as such. However, the character vectors of "4", "0", and "3" do not correspond to TRUE and FALSE logical vectors, thus outputting NA values.

as.logical(mixed3) converts all numeric vectors from the first concatenation string to logical vectors, and the lone character vector "3" to a logical vector. Since the TRUEs and FALSEs were converted to 1s and 0s and 1s and 0s do not correspond to logical vectors (instead, it's the other way around), NA values  were outputted since numbers cannot be converted to logical vectors. The character "3" vector also does not correspond to TRUE or FALSE, thus outputting an NA value.

### (c)
```{r Question 2c}
# convert all inputs to numeric vectors and then convert to logical vectors
as.logical(as.numeric(mixed3))
# any number other than 0 is outputted as TRUE
```

\newpage
## Question 3

### (a)
```{r Question 3a}
# seq() function
seq(1, 25)
# colon : operator
seq(1:25)
# seq_len() function
seq_len(25)
# seq_along function
x <- seq(1, 25)
seq_along(x)
# seq_along gives same length as object input, and by extension, same elements
```

### (b)
```{r Question 3b(i)}
# 'by' value is negative since the sequence is decreasing
seq(8, 2, -0.5)
```
```{r Question 3b(ii)}
# sequence from -1 to 5; omit elements, 3, 4, & 6, which correspond to 1, 2, & 4
rep(seq(-1, 5)[-c(3, 4, 6)], times = 5)
```

```{r Question 3b(iii)}
# use the TIMES parameter to repeat the entire sequence
rep(seq(5, 1), times = 5)
```

```{r Question 3b(iv)}
# use the EACH parameter to repeat individual numbers
rep(seq(5, 1), each = 5)
```

```{r Question 3b(v)}
# use the times 1:5 to repeat numbers in the sequence 1 to 5 times
# you can't use 'each' because the number of repetitions is not constant
rep(seq(10, 6), times = 1:5)
```

### (c)
The sequence cannot be generated using only the seq() and rep() functions because the boundaries of the sequence are changing each time, incrementally by 1, and are thus, not repeating in a consistent way.

```{r Question 3c}
# subtract 0 from first 5 numbers (10 to 6), then 1 from next 5, then 2, & so on
seq(10, 6, -1) - rep(0:4, each = 5)
```

\newpage
## Question 4
```{r Question 4}
k <- seq(0, 100, 1)
sin(k * pi)
```
The sine values for intervals of pi are outputted. Thinking about the unit circle, I know that for any whole number multiplied by pi, the sine output is 0. The results are what I would expect mathematically, because they are all near 0. The noticeable discrepancies are due to minor rounding and computing errors, particularly in dealing with irrational numbers like pi.

\newpage
## Question 5

### (a)
<!-- USE THE CODE CHUNK BELOW -->
```{r Question 5a}
# running times
x1 <- c(51, 40, 57, 34, 47, 50, 50, 56, 41, 38)

# find the coefficient of skewness

# numerator expression
n1 <- length(x1)
xbar1 <- mean(x1)
devs <- x1 - xbar1
cubed_devs <- devs^3
sum_cubed_devs <- sum(cubed_devs)
numerator1 <- sqrt(n1) * sum_cubed_devs
numerator1

# denominator expression
squared_devs <- devs^2
sum_squared_devs <- sum(squared_devs)
denominator1 <- sum_squared_devs^(3/2)
denominator1

# calculate skewness
skew1 <- numerator1 / denominator1
skew1
```

### (b)
```{r Question 5b}
# running times
x1 <- c(51, 40, 57, 34, 47, 50, 50, 56, 41, 38)

# find the coefficient of skewness using the one-pass formula

# numerator expression
n2 <- length(x1)
xbar2 <- mean(x1)
summations <- sum(x^3) - 3 * xbar2 * sum(x^2) + 2 * n2 * xbar2^3
numerator2 <- sqrt(n2) * summations
numerator2

# denominator expression
summation <- sum(x^2)
denominator2 <- summation - n2 * xbar2^2
denominator2

# calculate skewness
skew2 <- numerator2 / denominator2
skew2
```

### (c)
```{r Question 5c(a)}
# running times
x2 <- c(51, 40, 57, 34, 47, 50, 50, 56, 41, 38, 1e6)

# find the coefficient of skewness

# numerator expression
n1 <- length(x2)
xbar1 <- mean(x2)
devs <- x2 - xbar1
cubed_devs <- devs^3
sum_cubed_devs <- sum(cubed_devs)
numerator1 <- sqrt(n1) * sum_cubed_devs
numerator1

# denominator expression
squared_devs <- devs^2
sum_squared_devs <- sum(squared_devs)
denominator1 <- sum_squared_devs^(3/2)
denominator1

# calculate skewness
skew1 <- numerator1 / denominator1
skew1
```

```{r Question 5c(b)}
# running times
x1 <- c(51, 40, 57, 34, 47, 50, 50, 56, 41, 38, 1e6)

# find the coefficient of skewness using the one-pass formula

# numerator expression
n2 <- length(x1)
xbar2 <- mean(x1)
summations <- sum(x1^3) - 3 * xbar2 * sum(x1^2) + 2 * n2 * xbar2^3
numerator2 <- sqrt(n2) * summations
numerator2

# denominator expression
summation <- sum(x1^2)
denominator2 <- summation - n2 * xbar2^2
denominator2

# calculate skewness
skew2 <- numerator2 / denominator2
skew2
```

The results differ between from my answers in (a) and (b). They differ due to low precision errors. In particular, when the x values become too large, they are stored with low precision. However, in reality, they should be the same because adding a constant to a data set shifts the entire distribution, including the mean, but does not change the shape or skewness of the distribution.

### (d)
The results differed because of minor rounding and computing errors, particularly due to the data having very large numbers, such as 10^6.

\newpage
## Question 6

### (a)
```{r Question 6}
# Write code computing number of vector elements WITHOUT length() function. 
# Test your code on three vectors.

count_elements <- function(x) {
  # create a count elements function
  elements <- seq_len(1e6)
  # assign elements to a long numeric vector sequence
  valid_elements <- !is.na(x[elements])
  # recycle x in the sequence; turn all NAs to FALSE using !
  # if an element is NA, it IS NOT not a number, reporting FALSE
  sum(valid_elements)
  # use !is.na() instead of is.na() to report only TRUE values (values you want to count)
}
count_elements(x)
```

```{r Question 6a}
x <- numeric(0)
count_elements(x)
```

### (b)
<!-- USE THE CODE CHUNK BELOW -->
```{r Question 6b}
x <- seq(1, 99, 2)
count_elements(x)
```

### (c)
<!-- USE THE CODE CHUNK BELOW -->
```{r Question 6c}
x <- c("waffles", "friends", "work")
count_elements(x)
```

\newpage
## Question 7

### (a)
```{r Question 7}
# Write code reversing the order of vector elements in a vector WITHOUT rev() function. 
# Test your code on three vectors.

reverse_elements <- function(x){
  x[length(x):1] # output the sequence of x from length(x) to the first element
}

reverse_elements(x)
```

```{r Question 7a}
x <- numeric(0)
reverse_elements(x)
```

### (b)
<!-- USE THE CODE CHUNK BELOW -->
```{r Question 7b}
x <- seq(1, 99, 2)
reverse_elements(x)
```

### (c)
<!-- USE THE CODE CHUNK BELOW -->
```{r Question 7c}
x <- c("waffles", "friends", "work")
reverse_elements(x)
```

\newpage
## Question 8

### (a)
```{r Question 8ai}
first_elements <- function(x, n){
  x[seq_len(min(n, length(x)))] # output x using the seq_len function
  # seq_len can only have one input, so take the minimum of n or length of x
  # if n < length(x), compute the first n elements of x
  # if length(x) < n, compute the first length(x) elements of x
}

x <- numeric(0)
n <- 6
first_elements(x, n)
```

```{r Question 8a(ii)}
x <- seq(1, 99, 2)
n <- 6
first_elements(x, n)
```

```{r Question 8a(iii)}
x <- c("waffles", "friends", "work")
n <- 6
first_elements(x, n)
```

```{r Question 8a(iv)}
x <- numeric(0)
n <- 0
first_elements(x, n)
```

```{r Question 8a(v)}
x <- seq(1, 99, 2)
n <- 0
first_elements(x, n)
```

```{r Question 8(vi)}
x <- c("waffles", "friends", "work")
n <- 0
first_elements(x, n)
```

### (b)
```{r Question 8bi}
last_elements <- function(x, n){
  x[(max(length(x) - (n - 1), 1)):length(x)][length(x) > 1]
}

x <- numeric(0)
n <- 6
last_elements(x, n)
```

```{r Question 8b(ii)}
x <- seq(1, 99, 2)
n <- 6
last_elements(x, n)
```

```{r Question 8b(iii)}
x <- c("waffles", "friends", "work")
n <- 6
last_elements(x, n)
```

```{r Question 8b(iv)}
# verify with n = 0 that the code works on ONE of the aforementioned vectors
# see verified output for x <- numeric(0) below
x <- numeric(0)
n <- 0
last_elements(x, n)
```

```{r Question 8b(v)}
x <- seq(1, 99, 2)
n <- 0
last_elements(x, n)
```

```{r Question 8b(vi)}
x <- c("waffles", "friends", "work")
n <- 0
last_elements(x, n)
```
