---
title: "STATS 20: Week 2 Discussion Assignment"
author: "Brandon Liu"
date: "2025-01-15"
output:
  pdf_document:
    toc: true
    toc_depth: '4'
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
    code_folding: show
---

## PROBLEM 1
The Euclidean distance between two points is given by d equals the square root                  of the quantity of the change in x squared and the change in y squared.

Without using c() in your code, please write the code detailed below:

```{r question 1a}
# Please write code that calculates the Euclidean distance of numeric scalars.
# Scalars are x1, y1, x2, and y2.
# The values for these variables have been assigned.

x1 <- 1
y1 <- 1
x2 <- 1
y2 <- 2

d <- sqrt((x1 - x2)^2 + (y2 - y1)^2)
d
```

```{r question 1b}
# Write other code that calculates the Euclidean distance given inputs p1 & p2.
# p1 & p2 are numeric vectors with length = 2 of the form c(x1,y1) & c(x2,y2).
# The values of p1 and p2 have been assigned.
# I used d2 instead of d to specify the Euclidean distance for question 1b.

p1 <- c(1, 1)
p2 <- c(1, 2)

d2 <- sqrt((p2[1] - p1[1])^2 + (p2[2] - p1[2])^2)
d2
```

```{r question 1c}
# Write code that calculates the Euclidean distance given input pts.
# pts is a numeric vector with length = 4 of the form c(x1, y1, x2, y2).
# The value of pts has been assigned.
# I used d3 instead of d to specify the Euclidean distance for question 1c.

pts <- c(1, 1, 1, 2)

d3 <- sqrt((pts[3] - pts[1])^2 + (pts[4] - pts[2])^2)
d3
```

## PROBLEM 2
Given and using x <- letters[0:11], without using c(), output:

```{r question 2a}
# The elements at the odd indices of x "a" "c" "e" "g" "i" "k"
# index 0 is the same as index 1 for this sequence of letters

x <- letters[0:11]

x[seq(1, 11, by=2)]
```

```{r question 2b}
# The elements at the even indices of x: "b" "d" "f" "h" "j"

x[seq(0, 11, by=2)]
```

```{r question 2c}
# Every third element of x, starting at the third element of x: "c" "f" "i"

x[seq(0, 11, by=3)]
```

```{r question 2d}
# The first four elements of x in reverse order: "d" "c" "b" "a"

x[seq(4, 0)]
```

```{r question 2e}
# The second through fourth elements of x.
# Repeat the second once, the third twice, and the fourth thrice.
# For example, "b" "c" "c" "d" "d" "d"

x[rep(2:4, times = 1:3)]
```

```{r question 2f}
# The ninth element of x five times: "i" "i" "i" "i" "i"

x[seq(9, 9, length.out= 5)]
```

```{r question 2g}
# The empty character vector: character(0)

x[seq(length.out= 0)]
```

## PROBLEM 5
There are three kinematic equations for generally uniform linear motion.
They are: v = u + a*t, v^2 = u^2 + 2*a*d, and d = u*t + (a*t^2)/2 .

```{r question 5a}
# Write code that uses three numeric scalars u, a, and t.
# Compute v according to the relevant above kinematic equation.

# To solve for v, I assigned arbitrary integer values to u, a, and t.

u <- 3
a <- 1
t <- 2

v <- u + a*t
v
```

```{r question 5b}
# Write code that uses three numeric scalars u, a, and d.
# Compute v (not v^2) according to the relevant above kinematic equation.

# To solve for v, I assigned arbitrary integer values to u, a, and d.

u <- 3
a <- 1
d <- 4

# I transformed the v^2 equation into an equation for v.

v <- sqrt(u^2 + 2*a*d)
v
```

```{r question 5c}
# Write code that uses three numeric scalars u, a, and t.
# Compute d according to the relevant above kinematic equation.

# To solve for d, I assigned arbitrary integer values to u, a, and t.

u <- 3
a <- 1
t <- 2

d <- u*t + (a*t^2)/2
d
```

