---
title: "Stats 20: Week 5 Discussion Assignment"
author: "Brandon Liu"
date: "2025-02-04"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<!-- ## Concepts
** Logical Vectors

** Boolean algebra

** Logical operations in R

** Logical Indexing for Subsetting -->

#### Problem 1

Use `R` to identify the elements of the sequence $\text{NA}, 2^1, 2^2, \dots, 2^{15}$ that exceed the corresponding elements of the sequence $1^3, 2^3, \dots, 15^3, \text{NA}$.

```{r Problem 1}
seq1 <- c(NA, 2^(1:15))
seq2 <- c((1:15)^3, NA)

which(seq1 > seq2)
```

#### Problem 2

Let `c <- 3.141593`, and $f(x) = x\arcsin(\sin(c/x))$.

* Among the integers from 2 to 100, using `==`, `round(..., 6)`, find the integers that satisfy $f(x) \neq c$, respectively.
* Among the integers from 2 to 100, using `all.equal(..., ...)`, `all(... == ...)`, and `identical(..., ...)`, determine whether $f(x) = c$ for all $x$.

```{r Problem 2}
c <- 3.141593
sine_function <- 2:100 * asin(sin(c/2:100))
all.equal(sine_function, c)
all(sine_function == c)
identical(sine_function, c)
```

#### Problem 3

(1) Evaluation of a square root is achieved using the `sqrt()` function, but a warning will be issued when the argument is negative. Consider the following code which is designed to test whether a given value is positive before checking whether the square root of the value is less than 5.
```{r Problem 3}
testValue <- 7 
(testValue > 0) & (sqrt(testValue) < 5) 

testValue <- -7 
(testValue > 0) & (sqrt(testValue) < 5) 
```
Modify the code so that it continues to give the correct answer but without the warning.

```{r Problem 3 (1)}
testValue <- 7
(testValue > 0) & (sqrt(testValue) < 5) 

testValue <- -7
(testValue > 0) & (sqrt(!testValue) < 5)
```
(2)
Under what conditions would B need to be evaluated in the expression A || B ?
Answer: In this expression, B is only evaluated if A is FALSE. For example, if A \\ is TRUE, then B is not evaluated because the whole expression will be TRUE \\ regardless of the value of B. If A is FALSE, B will be evaluated.

#### Problem 4

(1) Using the values of a and b from this section, predict the output from each of these expressions, and then try them in `R`.
```{r Problem 4 (1), eval=FALSE}
a <- c(TRUE, FALSE, FALSE, TRUE)
b <- c(13, 7, 8, 2)

min(b) 
min(a) 
max(a) 
length(a) 
b * a
```

(2) What are the results of the following expressions?
```{r Problem 4 (2), eval=FALSE}
FALSE & c("STATS 20 is a great class!", "I love R!") 
# outputs an Error; c() function is not evaluating numeric or logical vectors
FALSE && c("STATS 20 is a great class!", "I love R!")
# outputs as FALSE; && only evaluates the first element.
TRUE && c("STATS 20 is a great class!", "I love R!")
# outputs an Error; c() function is not evaluating numeric or logical vectors
```

#### Problem 5
Given `scores <- c(85, 92, 76, 60, 45, 99, 88)`. Perform these operations:

(1) Find the number of scores that are greater than 80.
(2) Find the number of scores that are less than or equal to 60.
(3) Find the indices of all scores greater than 90 using `which()`.
(4) Extract all the scores that are multiples of 3.

```{r Problem 5 (1)}
scores <- c(85, 92, 76, 60, 45, 99, 88)
sum(scores > 80)
```
```{r Problem 5 (2)}
sum(scores <= 60)
```
```{r Problem 5 (3)}
which(scores > 90)
```
```{r Problem 5 (4)}
scores[scores %% 3 == 0]
```

### References
Braun, W. John, and Duncan J. Murdoch. A First Course in Statistical Programming with R. 3rd ed., Cambridge University Press, 2021, https://doi.org/10.1017/9781108993456. 